---
title: "HW4"
author: "Kailin"
date: "2024-10-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(DBI)
library(dbplyr)
library(bigrquery)
```


```{r}
project <- "hw-439518"
```

```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project
)
con
```
```{r}
dbListTables(con)
```

Write a first query that counts the number of rows of the `crime` table in the year 2016. Use code chunks with {sql connection = con} in order to write SQL code within the document.
```{sql connection = con}
SELECT count(primary_type), count(*)
FROM crime
WHERE year = 2016
LIMIT 10;
```

Next, count the number of arrests grouped by primary_type in 2016. Note that is a somewhat similar task as above, with some adjustments on which rows should be considered. Sort the results, i.e. list the number of arrests in a descending order.
```{sql connection = con}
SELECT primary_type, COUNT(*) AS arrest_count
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY primary_type
ORDER BY arrest_count DESC
LIMIT 10;
```

We can also use the date for grouping. Count the number of arrests grouped by hour of the day in 2016. You can extract the latter information from date via EXTRACT(HOUR FROM date). Which time of the day is associated with the most arrests?  
19:00 is associated with the most arrests.
```{sql connection = con}
SELECT EXTRACT(HOUR FROM date) AS hour_of_day, COUNT(*) AS arrest_count
FROM crime
WHERE year = 2016 AND arrest = TRUE
GROUP BY hour_of_day
ORDER BY arrest_count DESC
LIMIT 10;
```

Focus only on HOMICIDE and count the number of arrests for this incident type, grouped by year. List the results in descending order.
```{sql connection = con}
SELECT year, COUNT(*) AS arrest_count
FROM crime
WHERE primary_type = 'HOMICIDE' AND arrest = TRUE
GROUP BY year
ORDER BY arrest_count DESC
LIMIT 10;
```

Find out which districts have the highest numbers of arrests in 2015 and 2016. That is, count the number of arrests in 2015 and 2016, grouped by year and district. List the results in descending order.
District 11 has the highest numbers of arrests in 2015 and 2016.
```{sql connection = con}
SELECT year, district, COUNT(*) AS arrest_count
FROM crime
WHERE year IN (2015, 2016) AND arrest = TRUE
GROUP BY year, district
ORDER BY arrest_count DESC
LIMIT 10;
```

Lets switch to writing queries from within R via the DBI package. Create a query object that counts the number of arrests grouped by primary_type of district 11 in year 2016. The results should be displayed in descending order.
```{r}
query <- "SELECT primary_type, COUNT(*) AS arrest_count
          FROM crime
          WHERE year = 2016 AND district = 11 AND arrest = TRUE
          GROUP BY primary_type
          ORDER BY arrest_count DESC"
result1 <- dbGetQuery(con, query)
print(head(result1, 10))
```

Try to write the very same query, now using the dbplyr package. For this, you need to first map the crime table to a tibble object in R.
```{r}
cri <- tbl(con, "crime")
```

Again, count the number of arrests grouped by primary_type of district 11 in year 2016, now using dplyr syntax.
```{r}
query2 <- cri %>% select(primary_type, year, district, arrest) %>% 
  filter(year == 2016 & district == 11 & arrest == TRUE)%>%
  group_by(primary_type) %>% 
  summarise(total = n())%>%
  collect()
print(head(query2, 10))
```

Count the number of arrests grouped by primary_type and year, still only for district 11. Arrange the result by year.
```{r}
query3 <- cri %>% select(primary_type, year, district, arrest) %>%
  filter(district == 11 & arrest == TRUE) %>% 
  group_by(primary_type, year) %>% 
  summarise(total = n())%>%
  arrange(year) %>% 
  collect()
print(head(query3, 10))
```

```{r}
dbDisconnect(con)
```







